---
title: "Descriptive Statistics"
subtitle: 'Univariate Statistics'
author: "Martyna Ignaczak, Klaudia Gołubowska, Szymon Adamowicz"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown

---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(qwraps2_markup = "markdown")
library(qwraps2)
library(arsenal)
library(e1071)
library(haven)
library(papeR)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(summarytools)
library(classInt)
library(pastecs)
library(reporttools)
library(desctable)
library(psych)
library(frequency)
library(ggpubr)
library(ggforce)
library(ggdist)
library(gghalves)
library(gtsummary)
library(AER)

```

## Your turn!

Your task this week is to: prepare your own descriptive analysis for the
"CreditCard" dataset (AER package). It is a cross-sectional dataframe on
the credit history for a sample of applicants for a type of credit card.

## New dataset

this dataset displays info about credit cards of type "yes" and "no". based
on this test we will perform analisys of those two types.

```{r credit_cards}
data(CreditCard)

head(CreditCard, 10)

```

## Format
A data frame containing 1,319 observations on 12 variables.

card: Factor. Was the application for a credit card accepted?

reports: Number of major derogatory reports.

age: Age in years plus twelfths of a year.

income: Yearly income (in USD 10,000).

share: Ratio of monthly credit card expenditure to yearly income.

expenditure: Average monthly credit card expenditure.

owner: Factor. Does the individual own their home?

selfemp: Factor. Is the individual self-employed?

dependents: Number of dependents.

months: Months living at current address.

majorcards: Number of major credit cards held.

active: Number of active credit accounts.

## Age in relation to succsessful credit aplication

First analysis covers the relevance of age in getting approval for credit.
my first step will be to round the age to integer form so it is easier to
group it.

```{r}


# add column of age in integer form

age_analysis <- CreditCard
age_analysis$int_age <- as.integer(age_analysis$age)

# remove old column of age in double form

age_analysis <- subset(age_analysis, select = c (card, int_age))

# i found some age columns equal 0, it seems like wrong input/ missing data
# therefore i will get rid of them

age_analysis <- filter(age_analysis, int_age !=0)

age_analysis$age_group = as.character("")

age_analysis$age_group <- case_when(
  age_analysis$int_age >= 18 & age_analysis$int_age <= 25 ~ "18 - 25",
  age_analysis$int_age >= 26 & age_analysis$int_age <= 35 ~ "26 - 35",
  age_analysis$int_age >= 36 & age_analysis$int_age <= 50 ~ "36 - 50",
  age_analysis$int_age >= 51  ~ "51",

  TRUE ~ NA_character_ 
)

result <- age_analysis %>%
  group_by(age_group) %>%
  summarise(Yes_Count = sum(card == "yes"), No_Count = sum(card == "no"))

# Calculate the percentage of "Yes" statements for each group
result <- result %>%
  #group_by(age_group) %>%
  
  mutate(Percentage_Yes = (Yes_Count / (Yes_Count+No_Count)) * 100)

# View the result
print(result)

ggplot(age_analysis, aes(x = age_group, fill = card)) +
  geom_bar(position = "stack") +
  labs(title = "Contestants by Age Group and Decision",
       x = "Age Group",
       y = "Count") +
  scale_fill_manual(values = c("yes" = "blue", "no" = "red")) +
  theme_minimal()



```


Are the yearly incomes (in USD 10,000), credit card expenditures, age,
ratio of monthly credit card expenditure to yearly income -
significantly different for applicants for customers with different
credit risk ("card" variable - factor)?

Prepare a professional data visualizations, descriptive statistics'
tables and interpret them.

```{r my_summary_table}
# your code here
```
