---
title: "Descriptive Statistics"
subtitle: 'Univariate Statistics'
author: "Martyna Ignaczak, Klaudia Gołubowska, Szymon Adamowicz"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown

---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(qwraps2_markup = "markdown")
library(qwraps2)
library(arsenal)
library(e1071)
library(haven)
library(papeR)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(summarytools)
library(classInt)
library(pastecs)
library(reporttools)
library(desctable)
library(psych)
library(frequency)
library(ggpubr)
library(ggforce)
library(ggdist)
library(gghalves)
library(gtsummary)
library(AER)

```

## Your turn!

Your task this week is to: prepare your own descriptive analysis for the
"CreditCard" dataset (AER package). It is a cross-sectional dataframe on
the credit history for a sample of applicants for a type of credit card.

## New dataset

this dataset displays info about credit cards of type "yes" and "no". based
on this test we will perform analisys of those two types.

```{r credit_cards}
data(CreditCard)

head(CreditCard, 10)

```

```{r}







```
Are the yearly incomes (in USD 10,000), credit card expenditures, age,
ratio of monthly credit card expenditure to yearly income -
significantly different for applicants for customers with different
credit risk ("card" variable - factor)?

Prepare a professional data visualizations, descriptive statistics'
tables and interpret them.

```{r my_summary_table}
# your code here
```
