---
title: "Descriptive Statistics"
subtitle: 'Univariate Statistics'
author: "Martyna Ignaczak, Klaudia Gołubowska, Szymon Adamowicz"
date: "`r Sys.Date()`"
output: 
  rmdformats::readthedown

---

```{r setup1, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
options(qwraps2_markup = "markdown")
library(qwraps2)
library(arsenal)
library(e1071)
library(haven)
library(papeR)
library(dplyr)
library(tidyverse)
library(kableExtra)
library(summarytools)
library(classInt)
library(pastecs)
library(reporttools)
library(desctable)
library(psych)
library(frequency)
library(ggpubr)
library(ggforce)
library(ggdist)
library(gghalves)
library(gtsummary)
library(AER)

```

## Your turn!

Your task this week is to: prepare your own descriptive analysis for the
"CreditCard" dataset (AER package). It is a cross-sectional dataframe on
the credit history for a sample of applicants for a type of credit card.

## New dataset

this dataset displays info about credit cards of type "yes" and "no". based
on this test we will perform analisys of those two types.

```{r credit_cards}
data(CreditCard)

head(CreditCard, 10)

```

## Format
A data frame containing 1,319 observations on 12 variables.

card: Factor. Was the application for a credit card accepted?

reports: Number of major derogatory reports.

age: Age in years plus twelfths of a year.

income: Yearly income (in USD 10,000).

share: Ratio of monthly credit card expenditure to yearly income.

expenditure: Average monthly credit card expenditure.

owner: Factor. Does the individual own their home?

selfemp: Factor. Is the individual self-employed?

dependents: Number of dependents.

months: Months living at current address.

majorcards: Number of major credit cards held.

active: Number of active credit accounts.

## Age in relation to succsessful credit aplication

First analysis covers the relevance of age in getting approval for credit.
my first step will be to round the age to integer form so it is easier to
group it.

```{r}


# add column of age in integer form

age_analysis <- CreditCard
age_analysis$int_age <- as.integer(age_analysis$age)

# remove old column of age in double form

age_analysis <- subset(age_analysis, select = c (card, int_age))

# i found some age columns equal 0, it seems like wrong input/ missing data
# therefore i will get rid of them

age_analysis <- filter(age_analysis, int_age !=0)

age_analysis$age_group = as.character("")

age_analysis$age_group <- case_when(
  age_analysis$int_age >= 18 & age_analysis$int_age <= 25 ~ "18 - 25",
  age_analysis$int_age >= 26 & age_analysis$int_age <= 35 ~ "26 - 35",
  age_analysis$int_age >= 36 & age_analysis$int_age <= 50 ~ "36 - 50",
  age_analysis$int_age >= 51  ~ "51",

  TRUE ~ NA_character_ 
)

result <- age_analysis %>%
  group_by(age_group) %>%
  summarise(Yes_Count = sum(card == "yes"), No_Count = sum(card == "no"))

result <- result %>%
  mutate(Percentage_Yes = (Yes_Count / (Yes_Count+No_Count)) * 100)
```

The graph shows us that most of the samples were collected from age group 
26-35 years. Every age group had more instances of successful credit applications.
From this data we can assume that age isnt a significant factor in credit aplications.

```{r}

ggplot(age_analysis, aes(x = age_group, fill = card)) +
  geom_bar(position = "stack") +
  labs(title = "Accepted and denied credit cards in different age groups",
       x = "Age Group",
       y = "Ammount of applications") +
  scale_fill_manual(values = c("yes" = "lightgreen", "no" = "red")) +
  theme_minimal()



```
## Yearly income relevance in connection to application for credit

Now we can take a look at yearly income. 
```{r}

mean_income <- aggregate(income ~ card, data = CreditCard, FUN = mean)

ggplot(CreditCard, aes(x = income, colour = card, fill = card)) +
  geom_histogram(position = "stack", alpha=0.5, bins = 30, size = 0.75) +
  geom_vline(data = mean_income, aes(xintercept = income, color = card), linetype = "dashed", size=1) +
  scale_color_manual(values = c("yes" = "darkblue", "no" = "gold")) +
  scale_fill_manual(values = c("yes" = "darkblue", "no" = "gold")) +
  labs(title="Credit applications and their outcomes in relation to income",x = "Income in USD 10.000", y = "Applications") +
  theme_minimal()


```

From the histogram above we can determine that our income data doesn't 
follow normal distribution.

mean of accepted applications is slightly different from the denied ones
but the difference isn't really significant. The values mostly cover each
other and behave almost identically therefore i can't see any significant 
influence of income size to successful credit card application

## Next topic

opiszcie swoje wykresy zeby bylo to zaliczone jako analiza

```{r}

```

Are the yearly incomes (in USD 10,000), credit card expenditures, age,
ratio of monthly credit card expenditure to yearly income -
significantly different for applicants for customers with different
credit risk ("card" variable - factor)?

Prepare a professional data visualizations, descriptive statistics'
tables and interpret them.

```{r}
# your code here
```
