---
title: 'Descriptive Statistics'
subtitle: 'Bivariate Analysis'
date: "`r Sys.Date()`"
author: "Your name here"
output:
  html_document: 
    theme: cerulean
    highlight: textmate
    fontsize: 10pt
    toc: yes
    code_download: yes
    toc_float:
      collapsed: no
    df_print: default
    toc_depth: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup,	message = FALSE,	warning = FALSE,	include = FALSE}
library(dplyr)
library(tidyverse)
library(HSAUR3)
library(haven)
library(ggplot2)
library(gridExtra)
library(ppcor) # this package computes partial and semipartial correlations.
library(ltm) # this package computes point-biserial correlations.
library(devtools) 
install_github("markheckmann/ryouready") # please install package "ryouready" from github! (then # it)
library(ryouready) # this package computes nonlinear "eta" correlations.
library(GGally) # this package computes correlation matrix.
library(psych) # this package computes qualitative correlations.
library(DescTools) # this package computes qualitative correlations.
```


## Introduction

This is our first lab when we are considering 2 dimensions and instead of calculating univariate statistics by groups (or factors) of other variable - we will measure their common relationships based on co-variance and correlation coefficients. 

*Please be very careful when choosing the measure of correlation! In case of different measurument scales we have to recode one of the variables into weaker scale.

It would be nice to add some additional plots in the background. Feel free to add your own sections and use external packages.

## Data

This time we are going to use a typical credit scoring data with predefined "default" variables and personal demografic and income data. Please take a look closer at headers and descriptions of each variable.

```{r load-data, warning=TRUE, include=FALSE}
download.file("https://github.com/kflisikowski/ds/blob/master/bank_defaults.sav?raw=true", destfile ="bank_defaults.sav",mode="wb")
bank_defaults <- read_sav("bank_defaults.sav")
bank<-na.omit(bank_defaults)
bank$def<-as.factor(bank$default)
bank$educ<-as.factor(bank$ed)
```

## Scatterplots
We can add an estimated linear regression line:

```{r echo=FALSE, warning=TRUE}
# Basic scatter plot with the log of income
bank$logincome <- log(bank$income)
ggplot(data = bank, aes(x = age, y = logincome)) +
  geom_point(color = "black", size = 1) +
  geom_smooth(method=lm)+
  labs(title = "Age & Income",
       x = "Age",
       y = "Income")


```


```{r echo=FALSE, warning=TRUE}


cor(bank$logincome, bank$age)
pcor.test(bank$logincome, bank$age, bank$employ)


```

```{r echo=FALSE, warning=TRUE}
ggplot(data = bank, aes(x = educ, y = logincome)) +
  geom_boxplot(fill = "lightblue", color = "black", aes(group=ed)) +
  labs(title = "Education & income",
       x = "Education",
       y = "Income") +
  theme_minimal() 


```

```{r echo=FALSE, warning=TRUE}


cor(bank$logincome, bank$ed, method="kendall")


```


## Exercise 1. Contingency analysis.

Do you believe in the Afterlife?
https://nationalpost.com/news/canada/millennials-do-you-believe-in-life-after-life
A survey was conducted and a random sample of 1091 questionnaires is given in the form of the following contingency table:

```{r echo=FALSE, warning=FALSE}
x=c(435,147,375,134)
dim(x)=c(2,2)
dane<-as.table(x)
dimnames(dane)=list(Gender=c('Female','Male'),Believe=c('Yes','No'))
dane
fourfoldplot(dane)
```

Our task is to check if there is a significant relationship between the belief in the afterlife and gender. We can perform this procedure with the simple chi-square statistics and chosen qualitative correlation coefficient (two-way 2x2 table).

```{r echo=FALSE, warning=FALSE}
yes<-c(435,147)
no<-c(375,134)
#cohen.kappa(cbind(yes,no))
chisq.test(dane)
prop.table(dane)
```

As you can see we can calculate our chi-square statistic really quickly for two-way tables or larger. 
Now we can standardize this contingency measure to see if the relationship is significant.

```{r echo=FALSE, warning=FALSE}
Phi(dane)
#?ContCoef
#ContCoef(dane)
#CramerV(dane)
#TschuprowT(dane)
mosaicplot(dane)
barplot(dane)
```


## Exercise 2. Contingency analysis for the 'Titanic' data.

Let's consider the titanic dataset which contains a complete list of passengers and crew members on the RMS Titanic. It includes a variable indicating whether a person did survive the sinking of the RMS Titanic on April 15, 1912.
A data frame contains 2456 observations on 14 variables.

```{r load-data2, warning=TRUE, include=FALSE}
download.file("https://github.com/kflisikowski/ds/blob/master/titanic.csv?raw=true", destfile ="titanic.csv",mode="wb")
titanic <- read.csv("titanic.csv",row.names=1,sep=";")
```

The website http://www.encyclopedia-titanica.org/ offers detailed information about passengers and crew members on the RMS Titanic. According to the website 1317 passengers and 890 crew member were aboard.

8 musicians and 9 employees of the shipyard company are listed as passengers, but travelled with a free ticket, which is why they have NA values in fare. In addition to that, fare is truely missing for a few regular passengers. 

## Analisys
# 

```{r}
dane2<-titanic
cleanStatus <- dane2[!dane2$Status == "", ]

ggplot(cleanStatus, aes(x = Status)) +
  geom_bar() +
  labs(title = "Survivors and Victims",
       x = "Status",
       y = "Count") +
  theme_minimal()
```


