---
title: 'Descriptive Statistics'
subtitle: 'Bivariate Analysis'
date: "`r Sys.Date()`"
author: "Your name here"
output:
  html_document: 
    theme: cerulean
    highlight: textmate
    fontsize: 10pt
    toc: yes
    code_download: yes
    toc_float:
      collapsed: no
    df_print: default
    toc_depth: 5
editor_options: 
  markdown: 
    wrap: 72
---

```{r setup,	message = FALSE,	warning = FALSE,	include = FALSE}
library(dplyr)
library(tidyverse)
library(HSAUR3)
library(haven)
library(ggplot2)
library(gridExtra)
library(ppcor) # this package computes partial and semipartial correlations.
library(ltm) # this package computes point-biserial correlations.
library(devtools) 
install_github("markheckmann/ryouready") # please install package "ryouready" from github! (then # it)
library(ryouready) # this package computes nonlinear "eta" correlations.
library(GGally) # this package computes correlation matrix.
library(psych) # this package computes qualitative correlations.
library(DescTools) # this package computes qualitative correlations.
```


## Exercise 2. Contingency analysis for the 'Titanic' data.

Let's consider the titanic dataset which contains a complete list of passengers and crew members on the RMS Titanic. It includes a variable indicating whether a person did survive the sinking of the RMS Titanic on April 15, 1912.
A data frame contains 2456 observations on 14 variables.

```{r load-data2, warning=TRUE, include=FALSE}
download.file("https://github.com/kflisikowski/ds/blob/master/titanic.csv?raw=true", destfile ="titanic.csv",mode="wb")
titanic <- read.csv("titanic.csv",row.names=1,sep=";")
```

The website http://www.encyclopedia-titanica.org/ offers detailed information about passengers and crew members on the RMS Titanic. According to the website 1317 passengers and 890 crew member were aboard.

8 musicians and 9 employees of the shipyard company are listed as passengers, but travelled with a free ticket, which is why they have NA values in fare. In addition to that, fare is truely missing for a few regular passengers. 

## Analysis
# First we clean the empty values in Status column then create a geom bar with general data of survivours and victims

```{r}
dane2<-titanic
cleanStatus <- dane2[!dane2$Status == "", ]

ggplot(cleanStatus, aes(x = Status)) +
  geom_bar() +
  geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(title = "Survivors and Victims",
       x = "Status",
       y = "Count") +
  theme_minimal()


```

# Now lets look at gender in context of survival rates
```{r}

ggplot(cleanStatus, aes(x = Status, fill = Gender)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(title = "Survival Rate by Gender",
       x = "Status",
       y = "Count",
       fill = "Gender") +
  theme_minimal()


```

# corelation estimation
```{r}

cleanStatus$status_binary <- ifelse(cleanStatus$Status == "Survivor", 1, 0)
cleanStatus$gender_binary <- ifelse(cleanStatus$Gender == "Female", 1, 0)

status_gender <- table(cleanStatus$status_binary, cleanStatus$gender_binary)

phi_coefficient <- cor(cleanStatus$status_binary, cleanStatus$gender_binary)
print(phi_coefficient)


```
 0.47 result indicates a noticible relation between gender and survival rate. It is not an extremly strong
 relation but it's not insignificant either. This suggests that females had a higher likelihood of surviving 
compared to males, but other factors also contribute to survival chances.



# And now let's move on to vizualising the survival rates in terms of being a crew member or a passenger
```{r}
ggplot(cleanStatus, aes(x = Status, fill = Crew.or.Passenger.)) +
  geom_bar(position = "dodge") +
  geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(title = "Survival Rate by the Cause of the Excursion",
       x = "Status",
       y = "Crew Member or Passenger") +
  theme_minimal()

```
# corelation estimation
```{r}

cleanStatus$Crew.or.Passenger_binary <- ifelse(cleanStatus$Crew.or.Passenger == "Passenger", 1, 0)

status_Crew.or.Passenger <- table(cleanStatus$status_binary, cleanStatus$gender_binary)

phi_coefficient <- cor(cleanStatus$status_binary, cleanStatus$Crew.or.Passenger_binary)

print(phi_coefficient)


```
A correlation coefficient of 0.15 shows a weak positive relationship, meaning passengers were slightly more likely to survive than crew members on the Titanic. 

# Survival based on class

```{r}
ggplot(cleanStatus, aes(x = Class...Department, fill = Status)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(title = "Survival Rate by class and department of passengers",
       x = "Status",
       y = "Count",
       fill = "Class Department") 
```

 we single out only passengers of ship
 
```{r}

onlyPassengers <- cleanStatus[!cleanStatus$Crew.or.Passenger. == "Crew", ]

ggplot(onlyPassengers, aes(x = Class...Department, fill = Status)) +
  geom_bar(position = "dodge") +
  theme(axis.text.x = element_text(angle = 45, hjust = 1)) +
  geom_text(stat = "count", aes(label = ..count..), position = position_dodge(width = 0.9), vjust = -0.5) +
  labs(title = "Survival Rate by class and department of passengers",
       x = "Status",
       y = "Count",
       fill = "Class Department") 


```

```{r}





```


